---
title: "1.1: Introduction to RStudio"
---

## ⊳ Learning Outcomes

-   Students will be able to describe what computer programming languages (code) are, with R as an example.
-   Students will be able to describe the role of RStudio in programming with R.
-   Students will be able to describe the utility of each panel in RStudio.
-   Students will be able to perform basic math functions in the R console.

## ⊳ Lecture / Code-along

> ### Check-in (5 minutes)
>
> Let's all sign-up for RStudio!
>
> ### Rstudio Mini-tour
>
> Perform a live-coding mini-tour of RStudio.
>
> 1.  What do the different panels do?
> 2.  Console versus a script/Rmarkdown file
>
> ### Using R as a Calculator
>
> #### Using the Console
>
> You can do basic math in the Console (the bottom left part of the screen). The Console only understands R code, but it can also be treated as a calculator! Hence, we can type in some numbers and mathematical symbols.
>
> Try multiplying 5 and 3 in the Console (hint: \* means multiply). Hit `Enter` to run that line of code.
>
> #### R Markdown and Code Chunks
>
> R Markdown (.Rmd) is a file format that lets us incorporate text and code into one document seamlessly. In fact, it is the file format for this document!
>
> -   For writing text, you can type as you normally would.
> -   Code chunks are a bit different:
>     -   Near the top right of your screen you can toggle between viewing this document as under Source or Visual.
>     -   If you view this document under Source, you will see that all code chunks are sandwiched between " ```` ```{r} ```` " and " ```` ``` ```` ".
>     -   But under Visual, you can type R code in the lines under `{r}`.
>     -   To include text in chunks, you will need to put a `#` in front. R will not read anything in the line after `#` as code.
>
> Code chunks look like this:

```{r}
# This is a code chunk!
```

> A quick shortcut for adding a code chunk is `Ctrl + Alt + i` (`Cmd + Opt + i` on a Mac).
>
> Alternatively, you can go to Code \> Insert Chunk.
>
> To run a chunk of code, click the green arrow on the top right corner of the chunk.
>
> You can also run one or a few lines of code at a time by having your cursor on the line or highlighting multiple lines and hitting `Ctrl` + `Enter` (or `Cmd` + `Enter` on a Mac).
>
> #### Some Code Chunk Practice
>
> Let's work with an example code chunk.

```{r, message=FALSE, warning=FALSE}
# I am adding a comment to my code here.

432 - 11     # subtraction

12 + 18      # addition

8 / 4        # division

3 * 5        # multiplication
```

> Notice that if you run code in the Console rather than the code chunk in an R Markdown (.Rmd) file, you don't need to add the `{r}` to tell it that you are typing R code. The Console only understands R code anyway, so we don't need to tell it what it is!
>
> #### Math: Small Group Challenge
>
> Write a line of code to raise 2 to the 3rd power (often represented as $2^3$). This will likely require a little bit of Googling to figure out.
>
> First, write and run the code in the console.
>
> Next, write and run the code in the code chunk below. What did you have to do differently?

```{r, message=FALSE, warning=FALSE}
# 1) Console: 2^3 + Enter

# 2) Code chunk: 2^3 + Green arrow *OR* 2^3 + Ctrl + Shift

2^3
```
